---
title: "Introduction to mnis"
author: "Evan Odell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mnis Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`mnis` is an R package to pull data from the UK parliament through the Members' Name Information Service API. It emphasises simplicity and ease of use, so that users unfamiliar with APIs can easily retrieve large volumes of high quality data. 

All functions return data in JSON format, which is parsed to a list or data frame, depending on the data that is returned.

Functions are divided into three main types:

* Reference Functions

* Additional Information Functions

* Fixed Scope Functions

There is also the `mnis_base` function that allows for increased flexibility in all API requests.

##Installing mnis

###From CRAN

install.packages("mnis")

###From GitHub (Development Version)

install.packages("devtools")
devtools::install_github("EvanOdell/mnis")

###Load mnis

library(mnis)

##Using mnis

### mnis_Base function

The `mnis_Base()` function accepts requests to the mnis API. The full list of request options is available on the mnis website: http://data.parliament.uk/membersdataplatform/memberquery.aspx

### Reference Functions

A series of functions to return reference data. This data is useful for providing parameters for other function calls. These functions do not accept any arguments. 

`ref_AddressTypes()`

`ref_AnsweringBodies()`

`ref_Areas()`

`ref_AreaTypes()`

`ref_BiographyCategories()`

`ref_Cabinets()`

`ref_Committees()`

`ref_CommitteeTypes()`

`ref_Constituencies()`

`ref_ConstituencyAreas()`

`ref_ConstituencyTypes()`

`ref_Countries()`

`ref_Departments()`

`ref_DisqualificationTypes()`

`ref_Elections()`

`ref_ElectionTypes()`

`ref_EndReasons()`

`ref_ExperienceTypes()`

`ref_GovernmentPostDepartments()`

`ref_GovernmentPosts()`

`ref_GovernmentRanks()`

`ref_HonouraryPrefixes()`

`ref_HonourLists()`

`ref_Honours()`

`ref_InterestCategories()`

`ref_LordsMembershipTypes()`

`ref_LordsRanks()`

`ref_OppositionPostDepartments()`

`ref_OppositionPosts()`

`ref_OppositionRanks()`

`ref_OtherParliaments()`

`ref_ParliamentaryPosts()`

`ref_ParliamentaryRanks()`

`ref_ParliamentTypes()`

`ref_Parties()`

`ref_PartySubTypes()`

`ref_PhotoOutputs()`

`ref_Statuses()`

`ref_Titles()`

```{r eval=FALSE}
library(mnis)
x <- ref_Titles()
x

#   Title_Id         Name
#1        22        Canon
#2         1         Dame
#4         3          Hon
#3         2           Dr
#6         5         Lord
#5         4         Lady
#7         6       Lt Col
#8        21       Lt Gen
#9         7         Miss
#10        8           Mr
#11        9          Mrs
#12       10           Ms
#13       11         Prof
#14       19    Professor
#15       12          Rev
#16       13       Rev Dr
#17       18         Revd
#18       20      Revd Dr
#20       15          Sir
#19       14     Reverend
#21       16      The Rev
#22       17 The Reverend


```


###Additional Information Functions

`mnis_FullBiog()`

`mnis_BasicDetails()`

`mnis_BiographyEntries()`

`mnis_Committees()`

`mnis_Addresses()`

`mnis_Constituencies()`

`mnis_ElectionsContested()`

`mnis_Experiences()`

`mnis_GovernmentPosts()`

`mnis_Honours()`

`mnis_HouseMemberships()`

`mnis_Statuses()`

`mnis_Staff()`

`mnis_Interests()`

`mnis_KnownAs()`

`mnis_MaidenSpeeches()`

`mnis_OppositionPosts()`

`mnis_OtherParliaments()`

`mnis_ParliamentaryPosts()`

`mnis_Parties()`

`mnis_PreferredNames()`

#### mnis_Extra()

The `mnis_Extra()` function acts as a wrapper to the additional information functions. By default it calls all functions, which is equivalent to `mnis_FullBiog()`.

###Fixed Scope Functions

There are seven fixed scope functions that return ready-made datasets. They are: 

`mnis_PartyState()`

`mnis_ConstituencyResults()`

`mnis_Department()`

`mnis_GeneralElectionResults()`

`mnis_LordsType()`

`mnis_MemberDate()`

`mnis_PartyState()`

#### Fixed Scope Example

```{r eval=FALSE}
library(mnis)
x <- mnis_PartyState()
x

#    House Id                             Name MaleCount FemaleCount TotalCount
#  Commons  4                     Conservative       260          69        329
#  Commons 15                           Labour       130         101        231
#  Commons 29          Scottish National Party        36          18         54
#  Commons 17                 Liberal Democrat         8           1          9
#  Commons  7        Democratic Unionist Party         8           0          8
#  Commons  8                      Independent         1           3          4
#  Commons 30                        Sinn Fein         4           0          4
#  Commons 22                      Plaid Cymru         2           1          3
#  Commons 31 Social Democratic & Labour Party         2           1          3
#  Commons 38            Ulster Unionist Party         2           0          2
#  Commons 44                      Green Party         0           1          1
#  Commons 47                          Speaker         1           0          1
#  Commons 35            UK Independence Party         1           0          1

```




